Fundamental Problems with the Evidence Base of Adolescent Depression Treatments.

### 2024-01-06 apples_and_oranges.qmd
corrected title
changed plots slightly

---
title: Fundamental Problems with the Evidence Base of Adolescent Depression Treatments.
subtitle: Re-visiting the relative efficacy Control Conditions..
author:
  - name: Argyris Stringaris
    email: a.stringaris@ucl.ac.uk
    affiliations: 
        - id: ucl_1
          name: University College London
          department: Divisions of Psychiatry and Psychology and Language Science
          address: 1-19 Torrington Place
          city: London, UK
          state: 
          postal-code: WC1E 7HB
          country: United Kingdom
    attributes:
        corresponding: true
    note: 
  - name: Charlotte Burman
    email: charlotte.burman@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
  - name: Dayna Bhudia
    email: dayna.bhudia@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Lucy Foulkes
    email: lucy.foulkes@psych.ox.ac.uk
    affiliations:
        - id: OU
          name: University of Oxford
          department: Experimental Psychology
          address: Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Road
          city: Oxford
          postal-code: OX2 6GG
          country: United Kingdom
    note:
  - name: Carmen Moreno
    email: cmoreno@hggm.es
    affiliations:
        - id: HGM
          name: Hospital Gregorio Marañón
          department: Department of Psychiatry
          address: 46 C. del Dr. Esquerdo
          city: Madrid
          state:
          postal-code: 28007
          country: Spain
  - name: Samuele Cortese
    email: samuele.cortese@soton.ac.uk
    affiliations:
        - id: soton
          name: University of Southampton
          department: Centre for Innovation in Mental Health (CIMH), School of Psychology
          address: Highfield Campus, Building 44
          city: Southampton
          state:
          postal-code: SO17 1BJ
          country: United Kingdom
  - name: Georgina Krebs
    email: g.krebs@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note: 
abstract: |
  To follow
keywords: 
  - depression
  - adolescents
  - antidepressants
  - psychotherapy
date: last-modified
bibliography: emotion_concepts_paper.bib
format: 
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
---

@@

###############6. Function for plotting the SMDs
library(ggplot2)
library(stringr)

plot_means_function <- function(dataframe, subtitle) {
  dataframe$text_label <- 0

  for (i in 1:nrow(dataframe)) {
    dataframe$text_label[i] <- paste(
      "k = ", dataframe$n[i], "\n",
      round(dataframe$coef_means[i], 2),
      "[" ,
      round(dataframe$lower_ci[i], 2), 
      round(dataframe$upper_ci[i], 2),
      "]"
    ) 
  }

@@

  # Plotting
  plot_means <- ggplot(dataframe, aes(x = coef_means, y = 1:nrow(dataframe),
                                      colour = condition, label = text_label)) +
    geom_point(size = 3) +
    geom_errorbar(aes(xmin = lower_ci, xmax = upper_ci), width = 0.2, position = position_dodge(0.5)) +
    geom_text(vjust = +1.5, size = 4) +
    scale_size_continuous(guide = "none") +
    guides(colour = FALSE) + 
    scale_color_manual(values = c(redish_palette, blueish_palette)) +
    theme_minimal() +
    labs(x = "TE-random", y = NULL, title = "Adolescent Depression Trial Efficacy by Treatment Arm",
         subtitle = subtitle) +
    xlab("Standardized Mean Difference (SMD) with 95% CIs") +
    ylab("") +
    ylim(0, nrow(dataframe) + 1) +
    xlim(-3.0, 0.5) +
    geom_vline(xintercept = 0, linetype = "dashed", size = 1.5, colour = "grey") +
    geom_segment(x = -1.25, xend = -1.75, y = 4.7, yend = 4.7, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
    geom_text(x = -1.55, y = 4.9, label = "More Effective", color = "grey", vjust = 0.5, hjust = 1) +
    geom_segment(x = -1.35, xend = -0.85, y = 4.9, yend = 4.9, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
    geom_text(x = -1.05, y = 5.05, label = "Less Effective", color = "grey", vjust = 0.5, hjust = 0)+
    theme(axis.title.x = element_text(size = 14), 
          axis.text.x  = element_text(size = 14),
          axis.text.y = (element_blank()),
          plot.title = element_text(size = 16),
          plot.subtitle = element_text(size = 14)) +
    geom_curve(aes(x = 0.18, y = 4.73, xend = 0.02, yend = 4.5), color = "grey", curvature = -0.2, arrow = arrow(length = unit(0.25, "cm"), type = "closed")) +
    geom_text(x = 0.2, y = 4.75, label = "No Effect", color = "grey", vjust = 0.5, hjust = 0)

>>>>

---
title: Fundamental Problems with the Evidence Base of Adolescent Depression Treatments.
subtitle: Insights from the relative efficacy of Control Conditions.
author:
  - name: Argyris Stringaris
    email: a.stringaris@ucl.ac.uk
    affiliations: 
        - id: ucl_1
          name: University College London
          department: Divisions of Psychiatry and Psychology and Language Science
          address: 1-19 Torrington Place
          city: London, UK
          state: 
          postal-code: WC1E 7HB
          country: United Kingdom
    attributes:
        corresponding: true
    note: 
  - name: Charlotte Burman
    email: charlotte.burman@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
  - name: Dayna Bhudia
    email: dayna.bhudia@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Lucy Foulkes
    email: lucy.foulkes@psych.ox.ac.uk
    affiliations:
        - id: OU
          name: University of Oxford
          department: Experimental Psychology
          address: Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Road
          city: Oxford
          postal-code: OX2 6GG
          country: United Kingdom
    note:
  - name: Carmen Moreno
    email: cmoreno@hggm.es
    affiliations:
        - id: HGM
          name: Hospital Gregorio Marañón
          department: Department of Psychiatry
          address: 46 C. del Dr. Esquerdo
          city: Madrid
          state:
          postal-code: 28007
          country: Spain
  - name: Samuele Cortese
    email: samuele.cortese@soton.ac.uk
    affiliations:
        - id: soton
          name: University of Southampton
          department: Centre for Innovation in Mental Health (CIMH), School of Psychology
          address: Highfield Campus, Building 44
          city: Southampton
          state:
          postal-code: SO17 1BJ
          country: United Kingdom
  - name: Georgina Krebs
    email: g.krebs@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note: 
abstract: |
  To follow
keywords: 
  - depression
  - adolescents
  - antidepressants
  - psychotherapy
date: last-modified
bibliography: emotion_concepts_paper.bib
format: 
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
---

@@

###############6. Function for plotting the SMDs
plot_means_function <- function(dataframe, subtitle) {
  dataframe$text_label <- 0

  for (i in 1:nrow(dataframe)) {
    dataframe$text_label[i] <- paste(
      dataframe$condition[i], "\n",
      round(dataframe$coef_means[i], 2),
      "[" ,
      round(dataframe$lower_ci[i], 2), 
      round(dataframe$upper_ci[i], 2),
      "]"
    ) 
  }

@@

  # Plotting
  plot_means <- ggplot(dataframe, aes(x = coef_means, y = 1:nrow(dataframe),
                                      colour = condition, label = text_label)) +
    geom_point(label=dataframe$k) +
    geom_errorbar(aes(xmin = lower_ci, xmax = upper_ci), width = 0.2, position = position_dodge(0.5)) +
    geom_text(vjust = +1.5, size = 4) +
    scale_size_continuous(guide = "none") +
    guides(colour = FALSE) + 
    scale_color_manual(values = c(redish_palette, blueish_palette)) +
    theme_minimal() +
    labs(x = "TE-random", y = NULL, title = "Adolescent Depression Trial Efficacy by Treatment Arm",
         subtitle = subtitle) +
    xlab("Standardized Mean Difference (SMD) with 95% CIs") +
    ylab("") +
    ylim(0, nrow(dataframe) + 1) +
    xlim(-3.0, 0.5) +
    geom_vline(xintercept = 0, linetype = "dashed", size = 1.5, colour = "grey") +
    geom_segment(x = -1.25, xend = -1.75, y = 4.7, yend = 4.7, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
    geom_text(x = -1.55, y = 4.9, label = "More Effective", color = "grey", vjust = 0.5, hjust = 1) +
    geom_segment(x = -1.35, xend = -0.85, y = 4.9, yend = 4.9, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
    geom_text(x = -1.05, y = 5.05, label = "Less Effective", color = "grey", vjust = 0.5, hjust = 0)+
    theme(axis.title.x = element_text(size = 12), 
          axis.text.x  = element_text(size = 12),
          axis.text.y = (element_blank()),
          plot.title = element_text(size = 12),
          plot.subtitle = element_text(size = 12)) +
    geom_curve(aes(x = 0.18, y = 4.73, xend = 0.02, yend = 4.5), color = "grey", curvature = -0.2, arrow = arrow(length = unit(0.25, "cm"), type = "closed")) +
    geom_text(x = 0.2, y = 4.75, label = "No Effect", color = "grey", vjust = 0.5, hjust = 0)
